### 2013-01-07 documents/mixtures/response_to_referee.txt
started responding to referee

Dear Editor,

This was a very useful referee report and we have revised the paper
accordingly.  I think the referee comments have improved the paper
substantially.

David W. Hogg (on behalf of the authors)

-----

In what follows, the referee comments are indented and our responses
are not indented.

   1. Does the article contain significant new results and/or analysis
   and reflect sufficiently high scientific standards to warrant its
   publication in PASP?

   This paper gives a particularly lucid, concise and clear
   description of a relatively old idea. As such, it represents a
   significant contribution to the literature as a pedagogical
   reference. It should be published but see 4. below.

We agree with the referee that the idea is old but we don't agree with
any implication that the results shown are old.  All references we can
find in the literature show *how* to do mixture-of-Gaussian fitting in
real imaging, but don't give mixture-of-Gaussian approximations to
known profiles.  This paper, in that sense, is more of an "applied
math meets astronomy" paper than the previous work.  We will return to
this below, where we *do* make changes to the text in response to this
general concern of the referee.

   2. Is the paper written with the maximum conciseness compatible
   with accuracy and clarity?

   I thoroughly enjoyed reading the paper for its refreshing
   conciseness and clarity.

Thanks!

   3. Could the order or presentation or English be improved?  Editing
   of English and typographic errors will be corrected by the Journal
   copyediting staff and you do not need to specify these in your
   report.  But if there are more general problems with presentation
   or English please cite this in your report.

   Contractions (don't,  doesn't) should not be used.

   On page 8, paragraph 3: "If your PSF is not in MoG form ..." should read "If the
    PSF is not in MoG form ..."

   4. Do you have any comments or criticisms that may be helpful to the author(s)
   to improve or correct the paper? In cases that may be ambiguous, please specify
   whether you consider the suggested changes to be mandatory for publication or
   advisory.

   GENERAL COMMENT: Although I do think that this "Note" would stand on its own as
   a separate publication, I would nonetheless strongly suggest that the authors
   consider merging it into the forthcoming Lang et al. paper on The Tractor. This
   Note appears to be an "ad" for The Tractor. In the words of the authors, the main
   point of the Note is to emphasize the performance gain that the MoG approach
   would bring to the analysis of truly large galaxy datasets.  However, the current
   draft does not show any analysis of real data to support this performance claim.
   Other methods have been proposed in the past with similar promises of significant
   performance improvements (e.g., fast nonlinear regression via eigenimages,
   Brigham et al.). They looked promising on paper but were found to exhibit unexpected
   biases when applied to real data. I would hate to see this happen to MoGs. This
   is why it would be really nice to have the beautiful description of MoGs
   presented in this Note side-by-side with some tests on real data in the same paper.

   SPECIFIC COMMENTS:

   Page 2, paragraph 3: "There is no theoretical argument and only weak observational
   arguments that the rotation-supported parts of galaxies are always exponential
   and that the kinematically hot components are always more de Vaucouleurs-like".
   Although your warning about over-interpretating exp and dev components using
   bulges and disks is well taken (and others have given the same warning I might
   add), I think your statement above is too strong. On the theoretical side, some
   works (e.g., Brooks et al. 2011, ApJ, 728, 51; Agertz et al. 2011, MNRAS, 410,
   1391) have performed photometric decompositions of their model galaxies to see how
   photometric components mapped onto kinematical ones. The agreement is good and
   does suggest that it is often a reasonable mapping. It is certainly not a "radical"
   assumption to make. Also, while it is true that some dispersion-supported
   galaxies such as dwarf ellipticals have exponential profiles, they certainly do not
    make up the bulk of the galaxy population in large surveys. Finally, I note the
    recent ATLAS3D paper by Krajnovic et al. 2012, arXiv:1210.816 in which it was
   found that "early-type" galaxies with significant rotational support also
   exhibited significant exponential ("disk") components.

   Equations 2 and 3: I suggest you replace a_k by another symbol, c_k say, to
   avoid confusion with the \alpha's in the following equations.

   Page 6, first line: "The half-light inverse radius parameters \alpha^{exp} and \
   alpha^{dev} -- and the softening and cutoff radius parameters -- are taken from
   the SDSS codebase." I think you meant to write \alpha^{lux} and \alpha^{luv} here.

   Page 6, paragraph 2: "Although there is an analytic result for the dev profile,
   numerical integration of the concentrated profiles dev and luv to determine
   total fluxes can be challenging."  Why are you flagging the numerical integration of
    the dev profile as a challenge given that there is an analytic expression for it?

   Page 6, paragraph 2: "K-L divergence" --> "Kullback-Leibler divergence"

   Page 6 - Definition of "best": The authors should really have been "bolder" by
   insisting on the use of an information-theoretic measure here rather than the
   "traditional" chi-squared. The MoGs are a natural fit to an image entropy measure
   because maximizing entropy, say, would naturally lead to minimizing the number of
    components in a MoG. Chi-squared is also biased at low signal-to-noise ratios.
   Please consider.

   Page 6: Why is "badness" computed in a one-dimensional numerical integral? For
   the sake of computational speed? What happens when dealing with heavily blended
   objects? Where is the center for the 1D integral then? Please explain. A general
   explanation of how object deblending is performed with MoGs would be welcome.

   Page 7, paragraph 2: The statement "These -- M^lux = 6 and M^luv = 8 -- are good
    compromises between mixture complexity and quality of fit" warrants more
   explanation. Everytime I read about using MoGs for galaxy fitting, I have the same
   reaction: it needs a large number of highly redundant to model something a single
   exponential profile. Fans of the eigenimage decompositions are probably appalled!
    Granted the mathematical simplicity of MoGs can make up for this by enabling
   faster fitting times, but it is definitely an un-appealing aspect of the method.
   It would be helpful to comment on this to counter a reaction that I think may be
   shared by others. More importantly, I do not quite understand how you decided on
    M^lux = 6 and M^luv = 8. If I look at Table 2, the badness for lux has clearly
   converged and there is no gain in going higher than M=6. I cannot say the same
   thing for the luv MoG. Badness goes down by a factor of 16 going from M=6 to M=8
   and then goes down by a factor of 5 going from M=8 to M=10. How can you claim
   that your badness has "converged" for M=8 then? A quantitative threshold and its
   rationale are clearly needed here to justify the "good compromises" here. Please
   be more quantitative or clarify if I misunderstood something.

   Equation 19: I did not quite understand your distinction between the \Delta X
   vector here and the x vector in equation 2. Please clarify.

   ==============================================