### 2013-02-15 documents/mixtures/response_to_referee.txt
a few more referee points

Dear Editor,

This was a very useful referee report and we have revised the paper
accordingly.  I think the referee comments have improved the paper
substantially.

David W. Hogg (on behalf of the authors)

-----

We very much appreciate the thorough and sensible referee comments;
they have led to substantial improvements in the paper.  In what
follows, the referee comments are indented and our responses are not
indented.

   1. Does the article contain significant new results and/or analysis
   and reflect sufficiently high scientific standards to warrant its
   publication in PASP?

   This paper gives a particularly lucid, concise and clear
   description of a relatively old idea. As such, it represents a
   significant contribution to the literature as a pedagogical
   reference. It should be published but see 4. below.

We agree with the referee that the idea of using Gaussians to fit
galaxies is old but we don't agree with any implication that the
method or applied math results shown are old.  All references we can
find in the literature show *how* to do mixture-of-Gaussian fitting in
real imaging, but don't give mixture-of-Gaussian approximations to
known profiles.  This paper, in that sense, is more of an "applied
math meets astronomy" paper than the previous work.  We will return to
this below, but we *did* make changes to the text in response to this
general concern of the referee: We added sentences in the introductory
paragraphs and at the end of the paper to emphasize the difficulty of
rendering the profiles without these approximations.

To be more specific: we added some sentences to emphasize that the
currently used methods for rendering profiles are effectively mixtures
of delta functions (the profile is sampled at a finite set of
locations) and rendering requires as many PSF renderings as there are
sample points.  This is a huge computation.

We also added an introductory paragraph that points out that the code
complexity required to do rendering correctly is large and this leads
to engineering overheads in testing and correctness that go way beyond
the simple numerical speed losses.  We alsoo bring up the point that
conventional numerical integration methods effectively treat the
profiles as very complex mixtures of kernels.

[should we also add a late paragraph re-iterating and showing a comparison?]

   2. Is the paper written with the maximum conciseness compatible
   with accuracy and clarity?

   I thoroughly enjoyed reading the paper for its refreshing
   conciseness and clarity.

Thanks!

   3. Could the order or presentation or English be improved?  Editing
   of English and typographic errors will be corrected by the Journal
   copyediting staff and you do not need to specify these in your
   report.  But if there are more general problems with presentation
   or English please cite this in your report.

   Contractions (don't,  doesn't) should not be used.

Replaced all contractions.

   On page 8n, paragraph 3: "If your PSF is not in MoG form ..." should
   read "If the PSF is not in MoG form ..."

Fixed to "If the PSF is not represented in MoG form ...".

   4. Do you have any comments or criticisms that may be helpful to
   the author(s) to improve or correct the paper? In cases that may be
   ambiguous, please specify whether you consider the suggested
   changes to be mandatory for publication or advisory.

   GENERAL COMMENT: Although I do think that this "Note" would stand
   on its own as a separate publication, I would nonetheless strongly
   suggest that the authors consider merging it into the forthcoming
   Lang et al. paper on The Tractor. This Note appears to be an "ad"
   for The Tractor. In the words of the authors, the main point of the
   Note is to emphasize the performance gain that the MoG approach
   would bring to the analysis of truly large galaxy datasets.
   However, the current draft does not show any analysis of real data
   to support this performance claim.  Other methods have been
   proposed in the past with similar promises of significant
   performance improvements (e.g., fast nonlinear regression via
   eigenimages, Brigham et al.). They looked promising on paper but
   were found to exhibit unexpected biases when applied to real
   data. I would hate to see this happen to MoGs. This is why it would
   be really nice to have the beautiful description of MoGs presented
   in this Note side-by-side with some tests on real data in the same
   paper.

As noted above, we have added sentences in one paragraph and one new
paragraph that makes it clearer what the performance gains are.

[ should we also get quantitative at the end? ]

   SPECIFIC COMMENTS:

   Page 2, paragraph 3: "There is no theoretical argument and only
   weak observational arguments that the rotation-supported parts of
   galaxies are always exponential and that the kinematically hot
   components are always more de Vaucouleurs-like".  Although your
   warning about over-interpretating exp and dev components using
   bulges and disks is well taken (and others have given the same
   warning I might add), I think your statement above is too
   strong. On the theoretical side, some works (e.g., Brooks et
   al. 2011, ApJ, 728, 51; Agertz et al. 2011, MNRAS, 410, 1391) have
   performed photometric decompositions of their model galaxies to see
   how photometric components mapped onto kinematical ones. The
   agreement is good and does suggest that it is often a reasonable
   mapping. It is certainly not a "radical" assumption to make. Also,
   while it is true that some dispersion-supported galaxies such as
   dwarf ellipticals have exponential profiles, they certainly do not
   make up the bulk of the galaxy population in large
   surveys. Finally, I note the recent ATLAS3D paper by Krajnovic et
   al. 2012, arXiv:1210.816 in which it was found that "early-type"
   galaxies with significant rotational support also exhibited
   significant exponential ("disk") components.

Point well taken!  We softened this paragraph so that it no longer
makes any claim on this point; we now only attempt to induce some
feelings about conservativeness.

   Equations 2 and 3: I suggest you replace a_k by another symbol, c_k say, to
   avoid confusion with the \alpha's in the following equations.

Good point.  Fixed.

   Page 6, first line: "The half-light inverse radius parameters \alpha^{exp} and \
   alpha^{dev} -- and the softening and cutoff radius parameters -- are taken from
   the SDSS codebase." I think you meant to write \alpha^{lux} and \alpha^{luv} here.

Absolutely; fixed.

   Page 6, paragraph 2: "Although there is an analytic result for the dev profile,
   numerical integration of the concentrated profiles dev and luv to determine
   total fluxes can be challenging."  Why are you flagging the numerical integration of
    the dev profile as a challenge given that there is an analytic expression for it?

Indeed.  Adjusted that paragraph slightly and added a subsequent
paragraph explaining the implications for integrals over finite
domains.

   Page 6, paragraph 2: "K-L divergence" --> "Kullback-Leibler divergence"

fixed.

   Page 6 - Definition of "best": The authors should really have been "bolder" by
   insisting on the use of an information-theoretic measure here rather than the
   "traditional" chi-squared. The MoGs are a natural fit to an image entropy measure
   because maximizing entropy, say, would naturally lead to minimizing the number of
    components in a MoG. Chi-squared is also biased at low signal-to-noise ratios.
   Please consider.

   Page 6: Why is "badness" computed in a one-dimensional numerical integral? For
   the sake of computational speed? What happens when dealing with heavily blended
   objects? Where is the center for the 1D integral then? Please explain. A general
   explanation of how object deblending is performed with MoGs would be welcome.

All we are trying to do in this paper is come up with *replacements*
for the exponential and deVaucouleurs profiles and variants.  We are
not actually fitting real galaxy images.  Indeed that requires
deblending and so on.  Literally, all we are doing is replacing the
exp{r^{1/n}) with a sum of Gaussians.  We are not fitting real scenes.

[ we should make a text change here ]

   Page 7, paragraph 2: The statement "These -- M^lux = 6 and M^luv = 8 -- are good
    compromises between mixture complexity and quality of fit" warrants more
   explanation. Everytime I read about using MoGs for galaxy fitting, I have the same
   reaction: it needs a large number of highly redundant to model something a single
   exponential profile. Fans of the eigenimage decompositions are probably appalled!
    Granted the mathematical simplicity of MoGs can make up for this by enabling
   faster fitting times, but it is definitely an un-appealing aspect of the method.
   It would be helpful to comment on this to counter a reaction that I think may be
   shared by others. More importantly, I do not quite understand how you decided on
    M^lux = 6 and M^luv = 8. If I look at Table 2, the badness for lux has clearly
   converged and there is no gain in going higher than M=6. I cannot say the same
   thing for the luv MoG. Badness goes down by a factor of 16 going from M=6 to M=8
   and then goes down by a factor of 5 going from M=8 to M=10. How can you claim
   that your badness has "converged" for M=8 then? A quantitative threshold and its
   rationale are clearly needed here to justify the "good compromises" here. Please
   be more quantitative or clarify if I misunderstood something.

   Equation 19: I did not quite understand your distinction between the \Delta X
   vector here and the x vector in equation 2. Please clarify.

In other news, we also made the following changes to the text, in
response to other comments and issues:

- we did foo

- we did bar