### 2013-03-10 documents/mixtures/response_to_referee.txt
finished responding to referee issues

Dear Editor,

This was a very useful referee report and we have revised the paper
accordingly.  I think the referee comments have improved the paper
substantially.

David W. Hogg (on behalf of the authors)

-----

We very much appreciate the thorough and sensible referee comments;
they have led to substantial improvements in the paper.  In what
follows, the referee comments are indented and our responses are not
indented.

   1. Does the article contain significant new results and/or analysis
   and reflect sufficiently high scientific standards to warrant its
   publication in PASP?

   This paper gives a particularly lucid, concise and clear
   description of a relatively old idea. As such, it represents a
   significant contribution to the literature as a pedagogical
   reference. It should be published but see 4. below.

We agree with the referee that the idea of using Gaussians to fit
galaxies is old but we don't agree with any implication that the
method or applied math results shown are old.  All references we can
find in the literature show *how* to do mixture-of-Gaussian fitting in
real imaging, but don't give mixture-of-Gaussian approximations to
known profiles.  This paper, in that sense, is more of an "applied
math meets astronomy" paper than the previous work.  We will return to
this below, but we *did* make changes to the text in response to this
general concern of the referee: We added sentences in the introductory
paragraphs and at the end of the paper to emphasize the difficulty of
rendering the profiles without these approximations.

To be more specific: we added some sentences to emphasize that the
currently used methods for rendering profiles are effectively mixtures
of delta functions (the profile is sampled at a finite set of
locations) and rendering requires as many PSF renderings as there are
sample points.  This is a huge computation.

We also added an introductory paragraph that points out that the code
complexity required to do straight profile rendering (ie without the
MoGs) correctly is large and this leads to engineering overheads in
testing and correctness that go way beyond the simple numerical speed
losses.  We alsoo bring up the point that conventional numerical
integration methods effectively treat the profiles as very complex
mixtures of kernels.

   2. Is the paper written with the maximum conciseness compatible
   with accuracy and clarity?

   I thoroughly enjoyed reading the paper for its refreshing
   conciseness and clarity.

Thanks!

   3. Could the order or presentation or English be improved?  Editing
   of English and typographic errors will be corrected by the Journal
   copyediting staff and you do not need to specify these in your
   report.  But if there are more general problems with presentation
   or English please cite this in your report.

   Contractions (don't,  doesn't) should not be used.

Replaced all contractions.

   On page 8n, paragraph 3: "If your PSF is not in MoG form ..." should
   read "If the PSF is not in MoG form ..."

Fixed to "If the PSF is not represented in MoG form ...".

   4. Do you have any comments or criticisms that may be helpful to
   the author(s) to improve or correct the paper? In cases that may be
   ambiguous, please specify whether you consider the suggested
   changes to be mandatory for publication or advisory.

   GENERAL COMMENT: Although I do think that this "Note" would stand
   on its own as a separate publication, I would nonetheless strongly
   suggest that the authors consider merging it into the forthcoming
   Lang et al. paper on The Tractor. This Note appears to be an "ad"
   for The Tractor. In the words of the authors, the main point of the
   Note is to emphasize the performance gain that the MoG approach
   would bring to the analysis of truly large galaxy datasets.
   However, the current draft does not show any analysis of real data
   to support this performance claim.  Other methods have been
   proposed in the past with similar promises of significant
   performance improvements (e.g., fast nonlinear regression via
   eigenimages, Brigham et al.). They looked promising on paper but
   were found to exhibit unexpected biases when applied to real
   data. I would hate to see this happen to MoGs. This is why it would
   be really nice to have the beautiful description of MoGs presented
   in this Note side-by-side with some tests on real data in the same
   paper.

As noted above, we have added sentences in one paragraph and one new
paragraph that makes it clearer what the performance gains are.

   SPECIFIC COMMENTS:

   Page 2, paragraph 3: "There is no theoretical argument and only
   weak observational arguments that the rotation-supported parts of
   galaxies are always exponential and that the kinematically hot
   components are always more de Vaucouleurs-like".  Although your
   warning about over-interpretating exp and dev components using
   bulges and disks is well taken (and others have given the same
   warning I might add), I think your statement above is too
   strong. On the theoretical side, some works (e.g., Brooks et
   al. 2011, ApJ, 728, 51; Agertz et al. 2011, MNRAS, 410, 1391) have
   performed photometric decompositions of their model galaxies to see
   how photometric components mapped onto kinematical ones. The
   agreement is good and does suggest that it is often a reasonable
   mapping. It is certainly not a "radical" assumption to make. Also,
   while it is true that some dispersion-supported galaxies such as
   dwarf ellipticals have exponential profiles, they certainly do not
   make up the bulk of the galaxy population in large
   surveys. Finally, I note the recent ATLAS3D paper by Krajnovic et
   al. 2012, arXiv:1210.816 in which it was found that "early-type"
   galaxies with significant rotational support also exhibited
   significant exponential ("disk") components.

Point well taken!  We softened this paragraph so that it no longer
makes any claim on this point; we now only attempt to induce some
feelings about conservativeness.

   Equations 2 and 3: I suggest you replace a_k by another symbol, c_k
   say, to avoid confusion with the \alpha's in the following
   equations.

Good point.  Fixed.

   Page 6, first line: "The half-light inverse radius parameters
   \alpha^{exp} and \ alpha^{dev} -- and the softening and cutoff
   radius parameters -- are taken from the SDSS codebase." I think you
   meant to write \alpha^{lux} and \alpha^{luv} here.

Absolutely; fixed.

   Page 6, paragraph 2: "Although there is an analytic result for the
   dev profile, numerical integration of the concentrated profiles dev
   and luv to determine total fluxes can be challenging."  Why are you
   flagging the numerical integration of the dev profile as a
   challenge given that there is an analytic expression for it?

Indeed.  Adjusted that paragraph slightly and added a subsequent
paragraph explaining the implications for integrals over finite
domains.

   Page 6, paragraph 2: "K-L divergence" --> "Kullback-Leibler
   divergence"

fixed.

   Page 6 - Definition of "best": The authors should really have been
   "bolder" by insisting on the use of an information-theoretic
   measure here rather than the "traditional" chi-squared. The MoGs
   are a natural fit to an image entropy measure because maximizing
   entropy, say, would naturally lead to minimizing the number of
   components in a MoG. Chi-squared is also biased at low
   signal-to-noise ratios.  Please consider.

Chi-squared would definitely be biased if what we were fitting in
astronomical images were distributions of photon arrivals with no
background or read noise.  The presence of background (sky photons)
and read noise makes chi-squared a better statistic for optimization
than, say, the Poisson likelihood.  (That wouldn't be true if we were
fitting to *actual* images in which we know the sky level and read
noise, but we aren't.  Here we are just making *replacements* for the
standard profiles in a generic setting.)  All that said, the *precise*
measure you would use in *principle* depends (weakly) on the
properties of the data you have.  We added two sentences to this
effect in the relevant place (right after the K-L divergence mention).

As for the information-theoretic things mentioned by the referee: Yes,
in principle things like AIC and BIC and Bayes integral etc measures
for "how many components to use" would be very useful.  Unfortunately,
unlike with the decision about chi-squared vs Poisson likelihood, the
value of the AIC or BIC does depend very very sensitively on the noise
model (object, sky, and read noise levels) in every pixel of every
image.  That is, someone doing this correctly would use a higher M
expansion for the brighter galaxies and a lower M expansion for the
fainter galaxies.  That is very relevant and important.  We added text
about this where we give the M_exp=6 advice.  We note that although it
might make sense to go to large M from a data precision point of view,
it doesn't from a *model* precision point of view, since the models
don't fit the data at this level (there are spiral arms, deviations
from ellipsoidal symmetry and so on), as the referee evidently knows!
We added text about this too.

   Page 6: Why is "badness" computed in a one-dimensional numerical
   integral? For the sake of computational speed? What happens when
   dealing with heavily blended objects? Where is the center for the
   1D integral then? Please explain. A general explanation of how
   object deblending is performed with MoGs would be welcome.

All we are trying to do in this paper is come up with *replacements*
for the exponential and deVaucouleurs profiles and variants.  We are
not actually fitting real galaxy images.  Indeed that requires
deblending and so on.  Literally, all we are doing is replacing the
exp{r^{1/n}) with a sum of Gaussians.  We are not fitting real scenes.

In the Tractor papers which are forthcoming, we *are* solving this
deblending problem, but that is a separate matter, involving image
likelihoods, noise models, and priors (and utilities, perhaps) on
parameters, so it is way beyond the scope of this paper.  This paper
is just about the profiles *themselves*, not any specific data
analysis problem that we can solve with them.

In this paper *all* we are doing is the "applied math" problem of
replacing the analytic profiles with MoG approximations.  As we note
in the introductory paragraphs, *any* use of deV and exp profiles
involves some approximation when convolving with a realistic PSF.  We
are just arguing for using MoG approximations instead of the
alternatives, most of which are finite, adaptive samplings, which are
more computationally expensive and less accurate.

   Page 7, paragraph 2: The statement "These -- M^lux = 6 and M^luv =
   8 -- are good compromises between mixture complexity and quality of
   fit" warrants more explanation. Everytime I read about using MoGs
   for galaxy fitting, I have the same reaction: it needs a large
   number of highly redundant to model something a single exponential
   profile. Fans of the eigenimage decompositions are probably
   appalled!  Granted the mathematical simplicity of MoGs can make up
   for this by enabling faster fitting times, but it is definitely an
   un-appealing aspect of the method.  It would be helpful to comment
   on this to counter a reaction that I think may be shared by
   others. More importantly, I do not quite understand how you decided
   on M^lux = 6 and M^luv = 8. If I look at Table 2, the badness for
   lux has clearly converged and there is no gain in going higher than
   M=6. I cannot say the same thing for the luv MoG. Badness goes down
   by a factor of 16 going from M=6 to M=8 and then goes down by a
   factor of 5 going from M=8 to M=10. How can you claim that your
   badness has "converged" for M=8 then? A quantitative threshold and
   its rationale are clearly needed here to justify the "good
   compromises" here. Please be more quantitative or clarify if I
   misunderstood something.

We don't mean to claim that the badness has converged, and we modified
the text to say so.  We just mean that as a trade-off between code
speed and precision, in *practice* this is a good place to stop.
Again, this is just for *replacing* the analytic profiles, not for
general galaxy fitting.  For general galaxy fitting, we agree that
many more Gaussians can be required, especially for nearby
well-resolved galaxies.

   Equation 19: I did not quite understand your distinction between
   the \Delta X vector here and the x vector in equation 2. Please
   clarify.

Agreed.  We added three sentences shortly after equation (19)
explaining explicitly why we use \Delta x and not x here.  It relates
to the fact that the PSF is only used as a convolution kernel, never
as a model in itself.

In addition to the changes given above, we also made the following small changes:

[ and search email for other issues arising / suggestions / etc... ]

* We added an acknowledgement to the referee.